---
title: "transplantation_immport_data_download.R"
author: "Jieming Chen"
date: "June 23, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview
This script uses RImmPort to download structured, data from RImmPort, serialized them and then format the the RImmPort domains into tab-delimited files

***

### Part 1: download RImmPort-structured data (only need to do this once, so you can comment it out after one time)

1. It first sets up a tunnel from a shell script already in the system (Credit: Boris). Remember to close it later in the script.
```{r eval=FALSE}
system("/Users/jiemingchen/.start_tunnel.sh")
system("/Users/jiemingchen/.stop_tunnel.sh")
```

2. Connects to the ImmPort database in the Butte Lab (refer to this code: immport_connect.R)
```{r eval=FALSE}
source("/Users/jiemingchen/R_codes/immport_connect.R")
```

3. Serialize specific studies on local directory
```{r eval=FALSE}
serialzeStudyData(study_ids, data_dir=studies_dir)
```

### Part 2: Grab ImmPort data, structure them and store (serialize) them to a preset location

Part 2 of this script extracts domain information (23 domains for RImmPort). 
Note that this should be done after you split the studies into donors and recipients (or whichever stratification), otherwise it's going to compile everything into a single file
```{r}
library(RImmPort)
getListOfDomains()
```

1. Using a loop and write it to file; change the filename
```{r eval=FALSE}
for (i in 1:nrow(domains))
{
  data.donor = loadMultipleStudies(donor_ids,studies_dir,domains$`Domain Name`[i])
  write.table(data.donor, paste("/Users/jiemingchen/Documents/transplantation/",
                                domains$... = Code`[i],"_immport2tsv_donors_subjects_7studies.tsv", 
                                sep=""), 
              sep="\t", quote = FALSE, row.names = FALSE)
}
```