---
title: "R Notebook"
output: html_notebook
---

# library and function
```{r}
## my own library
source("/Users/jiemingchen/R_codes/jmRlib.R")

library(RImmPort)
library(DBI)
library(sqldf)
library(dplyr)
library(RMySQL)
library(ggplot2)
library(gridExtra)
library(gapminder)
library(plyr)
library(RColorBrewer)

## plot and save function
plot_age_sex <- function(data,flag,range) {
  names(data)[names(data)=="Patient_Sex"]  <- "Sex"
  names(data)[names(data)=="Donation_Age"]  <- "AGE"
  names(data)[names(data)=="Patient_Race2"]  <- "RACE"
  
  ## this is a histogram, age 40 is considered at bin 39; uses un-ddply a - nope
  # pmain = ggplot(data=data, aes(x=round(data$AGE), fill=Sex))
  # phisto = geom_histogram(breaks=seq(min(round(data$AGE)),round(max(data$AGE)),by=1), position = "dodge")
  
  ## this groups each age group as a category and then takes 15 min to plot!! uses ddply b -- nope
  # pmain = ggplot(b,aes(x=SEX,y=count,fill=Organ))
  # phisto = geom_bar(stat = "identity",color="white") + facet_wrap(~AGE,nrow=1)
  
  if(flag == 1)
  {
    ############
    # this one b1 
    pmain = ggplot(data=data, aes(x=AGE, y=count, fill=Sex))
    phisto = geom_bar(stat="identity", position = position_dodge())
    plabels = labs(x="Age",y="Count")
    pticks = scale_x_continuous(breaks=seq(min(round(data$AGE)),round(max(data$AGE)),by=1))
    paxes = theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
                  axis.title.y = element_text(face = "bold",colour = "black", size = 20),
                  axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15))
    ptitle = theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30))
    plegend = theme(legend.title = element_text(face = "bold", size = 15), legend.text = element_text(size = 14))
    
    myplot = pmain + phisto + plabels + pticks + paxes + ptitle + plegend
    return(myplot)
  } else if(flag == 2)
  {
    ########
    ## this plots each x axis as an interaction of age and sex and then adjust label for age
    ## order of Sex and AGE matter - uses b
    ## plots organ with this data
    pmain = ggplot(data, aes(x = as.numeric(interaction(Sex,AGE)), y=count, fill=ORGAN))
    phisto = geom_bar(stat = "identity") 
    plabels = labs(x="Age",y="Count")
    # pticks = scale_x_continuous(breaks=range, labels=seq(min(round(data$AGE)),round(max(data$AGE)),by=1))
    #pticks = scale_x_continuous(breaks=range,labels=seq(min(round(data$AGE)),round(max(data$AGE)),by=1))
    # pticks = scale_x_continuous() ##debug
    paxes = theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
                  axis.title.y = element_text(face = "bold",colour = "black", size = 20),
                  axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15))
    ptitle = theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30))
    plegend = theme(legend.title = element_text(face = "bold", size = 15), legend.text = element_text(size = 14))
    
    pcolor = scale_fill_manual(values=c("#00BA38","#F8766D","#619CFF")) ## only for donors
    myplot = pmain + phisto + plabels + paxes + ptitle + plegend + pcolor
    return(myplot)
  } else
  {
    ########
    ## this plots each x axis as an interaction of age and sex and then adjust label for age
    ## order of Sex and AGE matter - uses b
    ## plots study with this data
    pmain = ggplot(data, aes(x = as.numeric(interaction(Sex,AGE)), y=count, fill=RACE))
    phisto = geom_bar(stat = "identity") 
    plabels = labs(x="Age",y="Count")
    # pticks = scale_x_continuous(breaks=range, labels=seq(min(round(data$AGE)),round(max(data$AGE)),by=1))
    #pticks = scale_x_continuous(breaks=range,labels=seq(min(round(data$AGE)),round(max(data$AGE)),by=1))
    # pticks = scale_x_continuous() ##debug
    paxes = theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
                  axis.title.y = element_text(face = "bold",colour = "black", size = 20),
                  axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15))
    ptitle = theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30))
    plegend = theme(legend.title = element_text(face = "bold", size = 15), legend.text = element_text(size = 14))
    
    # pcolor = scale_fill_manual(values=c("#00BA38","#F8766D","#619CFF")) ## only for donors
    myplot = pmain + phisto + plabels + paxes + ptitle + plegend
    return(myplot)
  }
}
```


#------ COMPARING IMMPORT, EHR and UNOS/OPTN ------ 
## input data
```{r}
##-----------------------------------------------------------
## input data
# extract only kidney
# fsieve -s /Users/jiemingchen/Documents/transplantation/a_donor/immport/finals.iTx/final.iTx.DM.v15.nonkidneyLDs.txt combined_emr_immport_unos-all-updated-stand.txt -m 
idata <- read.table("/Users/jiemingchen/Documents/transplantation/a_donor/combined/sieved-combined_emr_immport_unos-all-updated-stand.txt", header = T, sep = "\t")
a.d <- idata[!is.na(idata$Donation_Age) & !is.na(idata$Sex),]





b3.dd <- mutate(a.d, source_sex = paste(Data_Source, "_", Sex, sep = ""))
b3.dd <- ddply(.data=b3.dd, .variables="Data_Source", .fun=transform, sum.n = length(Data_Source))
b3.dd <- ddply(.data=b3.dd, .variables="source_sex", .fun=transform, sum.sex.n = length(source_sex),
                                                                     median.sex.n = median(Donation_Age))

b3.d1 <- ddply(.data = b3.dd, .variables = c("Sex","Data_Source"), .fun=summarise, 
           count = length(Data_Source),
           proportion = count/sum.n[1])

b3.d2 <- ddply(.data = b3.dd, .variables = c("Donation_Age","Sex","Data_Source"), .fun=summarise, 
           count = length(Donation_Age),
           proportion = count/sum.n[1],
           proportion.sex = count/sum.sex.n[1])
```


## gender
### proportion plot
```{r}

b3.d1 %>%
  ggplot(aes(Sex, proportion, fill = Sex)) +
  geom_bar(stat = "identity") +
  facet_grid(Data_Source ~., scale = "fixed") + 
  theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
              axis.title.y = element_text(face = "bold",colour = "black", size = 20),
              axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15)) +
  theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30)) + 
  scale_fill_brewer(palette = "Set1")

```

### line plot
```{r}
## PLOT split by dataset; plot by age and sex using ggplot2; line plot proportion by total in dataset
b3.d2 %>% 
  ggplot(aes(Donation_Age, proportion, color = Sex)) +
  # ggplot(aes(Donation_Age, proportion.sex, color = Sex)) +
  geom_line(size=1.5) +
  facet_grid(Data_Source ~., scale = "free_y") +
  labs(x="Age",y="Proportion") +
  scale_x_continuous(breaks=seq(min(b3.d2$Donation_Age),max(b3.d2$Donation_Age),by=5)) +
  theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
        axis.title.y = element_text(face = "bold",colour = "black", size = 20),
        axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15)) +
  theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30)) +
  theme(legend.title = element_text(face = "bold", size = 15), legend.text = element_text(size = 14)) +
  geom_vline(xintercept = b3.dd$median.sex.n, color = "red") + ## use illustrator to do this individually
  scale_color_brewer(palette = "Set1")

## medians to check
unique(b3.dd %>% select(source_sex, median.sex.n))
median(b3.dd[b3.dd$Sex == "Female" & b3.dd$Data_Source == "ImmPort_DR19",]$Donation_Age)
median(b3.dd[b3.dd$Sex == "Male" & b3.dd$Data_Source == "ImmPort_DR19",]$Donation_Age)
median(b3.dd[b3.dd$Sex == "Female" & b3.dd$Data_Source == "UCSF_EMR",]$Donation_Age)
median(b3.dd[b3.dd$Sex == "Male" & b3.dd$Data_Source == "UCSF_EMR",]$Donation_Age)
median(b3.dd[b3.dd$Sex == "Female" & b3.dd$Data_Source == "UNOS_OPTN",]$Donation_Age)
median(b3.dd[b3.dd$Sex == "Male" & b3.dd$Data_Source == "UNOS_OPTN",]$Donation_Age)
median(b3.dd[b3.dd$Data_Source == "ImmPort_DR19",]$Donation_Age)

```

### bar plot
```{r}
# ## PLOT split by dataset; plot by age and sex using ggplot2; BARS
# pmain <- ggplot(b3.d2, aes(Donation_Age, count, fill = Sex))
# # pmain <- ggplot(b3.d[b3.d$Sex != "Not_Specified", ], aes(Donation_Age, count, fill = Sex)) ## remove unknown sex
# phisto <- geom_bar(stat = "identity", position=position_dodge())
# pfacet <- facet_grid(Data_Source ~., scale = "free_y") 
# plabels <- labs(x="Age",y="Count")
# pticks <- scale_x_continuous(breaks=seq(min(b3.d2$Donation_Age),max(b3.d2$Donation_Age),by=1))
# paxes <- theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
#               axis.title.y = element_text(face = "bold",colour = "black", size = 20),
#               axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15))
# ptitle <- theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30))
# plegend <- theme(legend.title = element_text(face = "bold", size = 15), legend.text = element_text(size = 14))
# pmain + phisto + pfacet + plabels + pticks + paxes + ptitle + plegend + scale_fill_brewer(palette = "Set1")
# # savePlot("plot_combined_immport_emr_unos-jun2016_age_sex_kidney_donors_bars.png", type = "png")

```


### stats test -- use KS test
```{r}
immport.b3 <- b3.d2[b3.d2$Data_Source == "ImmPort_DR19",]
imm.t <- sum(immport.b3$count)
unosopt.b3 <- b3.d2[b3.d2$Data_Source == "UNOS_OPTN",]
unos.t <- sum(unosopt.b3$count)
ucsfemr.b3 <- b3.d2[b3.d2$Data_Source == "UCSF_EMR",]
emr.t <- sum(ucsfemr.b3$count)
```

### within immport
```{r}
imm.f <- immport.b3[immport.b3$Sex=="Female",]
imm.m <- immport.b3[immport.b3$Sex=="Male",]

a <- merge(imm.f, imm.m, by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] = 0

prob.x <- a$count.x/imm.t
prob.y <- a$count.y/imm.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

```

#### merge immport and unos
```{r}
## FEMALE
a <- merge(unosopt.b3[unosopt.b3$Sex=="Female",], immport.b3[immport.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] = 0

## chi sq test, expected = unos
# z <- a[a$count.x>20 & a$count.y>20,]
# prob.y <- z$count.y/sum(z$count.y)
# chisq.test(z$count.x, prob.y) ## x can be count data, but y needs to be proportion NOT counts

## KS test
prob.x <- a$count.x/unos.t
prob.y <- a$count.y/imm.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

## qqplot
# qqplot(prob.x, prob.y)
# abline(0,1, col="red", lty=3)


## MALE
a <- merge(unosopt.b3[unosopt.b3$Sex=="Male",], immport.b3[immport.b3$Sex=="Male",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] <- 0

## chi sq test, expected = unos
# z <- a[a$count.x>20 & a$count.y>20,]
# prob.y <- z$count.y/sum(z$count.y)
# chisq.test(z$count.x, prob.y) ## x can be count data, but y needs to be proportion NOT counts

## KS test
## test for heterogeneity - so it doesnt test for similarity 
prob.x <- a$count.x/unos.t
prob.y <- a$count.y/imm.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

## qqplot
# qqplot(prob.x, prob.y)
# abline(0,1, col="red", lty=3)


#######
## fishers between male and female
# a <- merge(immport.b3[immport.b3$Sex=="Male",], immport.b3[immport.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- a[,c("Donation_Age","count.x","count.y")]
# names(a) = c("Donation_Age","male.immport","female.immport")
# a <- merge(a, unosopt.b3[unosopt.b3$Sex=="Male",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- merge(a, unosopt.b3[unosopt.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- a[,c("Donation_Age","male.immport","female.immport","count.x","count.y")]
# names(a) = c("Donation_Age","male.immport","female.immport","male.unos","female.unos")
# a[is.na(a)] <- 0
# 
# fish <- function(x) {
#   aa=
#   bb=cbind(aa$p.value,aa$estimate)
#   return(bb)
# }
# 
# ft.r.p <- as.data.frame(apply(a,1,function(x) fisher.test(matrix(x[2:5],2,2), alternative="two.sided")$p.value))
# ft.r.e <- as.data.frame(apply(a,1,function(x) fisher.test(matrix(x[2:5],2,2), alternative="two.sided")$estimate))
# ft.r <- cbind(a$Donation_Age,ft.r.p,ft.r.e); names(ft.r) <- c("Donation_Age","fish.p","fish.OR_est")
# ft.r$fish.p.bon <- as.data.frame(apply(as.matrix(ft.r$fish.p),1,function(x) min(x*nrow(ft.r),1)) )
```

#### merge emr and immport
```{r}
## FEMALE
a <- merge(ucsfemr.b3[ucsfemr.b3$Sex=="Female",], immport.b3[immport.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] <- 0

## chi sq test, expected = unos
# z <- a
# prob.y <- z$count.y/sum(z$count.y)
# chisq.test(z$count.x, prob.y) ## x can be count data, but y needs to be proportion NOT counts

## KS test
prob.x <- a$count.x/emr.t
prob.y <- a$count.y/imm.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

## qqplot
# qqplot(prob.x, prob.y)
# abline(0,1, col="red", lty=3)

#
## MALE
a <- merge(ucsfemr.b3[ucsfemr.b3$Sex=="Male",], immport.b3[immport.b3$Sex=="Male",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] = 0

## chi sq test, expected = unos
# z <- a[a$count.x>20 & a$count.y>20,]
# z <- a
# prob.y <- z$count.y/sum(z$count.y)
# chisq.test(z$count.x, prob.y) ## x can be count data, but y needs to be proportion NOT counts

## KS test
prob.x <- a$count.x/emr.t
prob.y <- a$count.y/imm.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

## qqplot
# qqplot(prob.x, prob.y)
# abline(0,1, col="red", lty=3)

#


#
## fishers between male and female
# a <- merge(ucsfemr.b3[ucsfemr.b3$Sex=="Male",], immport.b3[immport.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- a[,c("Donation_Age","count.x","count.y")]
# names(a) = c("Donation_Age","male.immport","female.immport")
# a <- merge(a, ucsfemr.b3[ucsfemr.b3$Sex=="Male",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- merge(a, ucsfemr.b3[ucsfemr.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- a[,c("Donation_Age","male.immport","female.immport","count.x","count.y")]
# names(a) <- c("Donation_Age","male.immport","female.immport","male.emr","female.emr")
# a[is.na(a)] = 0
# 
# fish <- function(x) {
#   aa=
#     bb <- cbind(aa$p.value,aa$estimate)
#   return(bb)
# }
# 
# ft.r.p <- as.data.frame(apply(a,1,function(x) fisher.test(matrix(x[2:5],2,2), alternative="two.sided")$p.value))
# ft.r.e <- as.data.frame(apply(a,1,function(x) fisher.test(matrix(x[2:5],2,2), alternative="two.sided")$estimate))
# ft.r <- cbind(a$Donation_Age,ft.r.p,ft.r.e); names(ft.r) <- c("Donation_Age","fish.p","fish.OR_est")
# ft.r$fish.p.bon = as.data.frame(apply(as.matrix(ft.r$fish.p),1,function(x) min(x*nrow(ft.r),1)) )
```

#### merge emr and unos
```{r}
############################
## #### 
#
## FEMALE
a <- merge(ucsfemr.b3[ucsfemr.b3$Sex=="Female",], unosopt.b3[unosopt.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] <- 0

## chi sq test, expected = unos
# z <- a
# prob.y <- z$count.y/sum(z$count.y)
# chisq.test(z$count.x, prob.y) ## x can be count data, but y needs to be proportion NOT counts

## KS test
prob.x <- a$count.x/emr.t
prob.y <- a$count.y/unos.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

## qqplot
# qqplot(prob.x, prob.y)
# abline(0,1, col="red", lty=3)


## MALE
a <- merge(ucsfemr.b3[ucsfemr.b3$Sex=="Male",], unosopt.b3[unosopt.b3$Sex=="Male",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] <- 0

## chi sq test, expected = unos
# z <- a[a$count.x>20 & a$count.y>20,]
# z <- a
# prob.y <- z$count.y/sum(z$count.y)
# chisq.test(z$count.x, prob.y) ## x can be count data, but y needs to be proportion NOT counts

## KS test
prob.x <- a$count.x/emr.t
prob.y <- a$count.y/unos.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

## qqplot
# qqplot(prob.x, prob.y)
# abline(0,1, col="red", lty=3)



#
## fishers between male and female
# a <- merge(ucsfemr.b3[ucsfemr.b3$Sex=="Male",], ucsfemr.b3[ucsfemr.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- a[,c("Donation_Age","count.x","count.y")]
# names(a) = c("Donation_Age","male.emr","female.emr")
# a <- merge(a, unosopt.b3[unosopt.b3$Sex=="Male",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- merge(a, unosopt.b3[unosopt.b3$Sex=="Female",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
# a <- a[,c("Donation_Age","male.emr","female.emr","count.x","count.y")]
# names(a) <- c("Donation_Age","male.emr","female.emr","male.unos","female.unos")
# a[is.na(a)] <- 0
# 
# fish <- function(x) {
#   aa=
#     bb=cbind(aa$p.value,aa$estimate)
#   return(bb)
# }
# 
# ft.r.p <- as.data.frame(apply(a,1,function(x) fisher.test(matrix(x[2:5],2,2), alternative="two.sided")$p.value))
# ft.r.e <- as.data.frame(apply(a,1,function(x) fisher.test(matrix(x[2:5],2,2), alternative="two.sided")$estimate))
# ft.r <- cbind(a$Donation_Age,ft.r.p,ft.r.e); names(ft.r) = c("Donation_Age","fish.p","fish.OR_est")
# ft.r$fish.p.bon <- as.data.frame(apply(as.matrix(ft.r$fish.p),1,function(x) min(x*nrow(ft.r),1)) )
```

## race/ethnicity
### input
```{r}
## remove missing Race
## remove hispanics because UNOS has hispanic/latinos as a "race" category, which is inconsistent with the ImmPort clinical trial data
## remove hispanics from b3.dd first so that the denominator (total) for proportion is calculated correctly
a.r <- idata %>% filter(!is.na(Donation_Age)) %>%
                 filter(!is.na(Race))
                 # filter(Race != "Hispanic/Latino")
b4.rr <- ddply(.data=a.r, .variables="Data_Source", .fun=transform, sum.n = length(Data_Source))

## calculate proportion
b4.r1 <- ddply(.data = b4.rr, .variables = c("Donation_Age","Race","Data_Source"), .fun=summarise, count = length(Donation_Age), proportion = count/sum.n[1])

b4.r2 <- ddply(.data = b4.rr, .variables = c("Donation_Age","Sex","Race","Data_Source"), .fun=summarise, count = length(Donation_Age), proportion = count/sum.n[1])

```

## PLOT split by dataset; plot by ETHNICITY using ggplot2; bars
```{r}
## get colors
getPalette <- brewer.pal(9, "Set1")

## plot
b4.r1 %>% 
  ggplot(aes(x=Donation_Age, y=proportion, fill=Race)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) + 
  facet_grid(Data_Source ~., scale = "free_y") + 
  labs(x="Age of Donation",y="Proportion") +
  theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
              axis.title.y = element_text(face = "bold",colour = "black", size = 20),
              axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15)) +
  theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30)) +
  theme(legend.title = element_text(face = "bold", size = 15), legend.text = element_text(size = 14)) + 
  scale_fill_brewer(palette = "Set1") +
  scale_x_continuous(breaks=seq(0,80,by=5))
# ggsave("barplots-gender-aggregate-immport_emr_unos.pdf", device = "pdf")

## stacked bar split by gender and stacked by ethnicity
b4.r2 %>%
  ggplot(aes(Sex, proportion, fill = Race)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  facet_grid(Data_Source ~., scale = "fixed") + 
  theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
              axis.title.y = element_text(face = "bold",colour = "black", size = 20),
              axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15)) +
  theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30)) + 
  scale_fill_brewer(palette = "Set1")
```

## stat test for similarity; KS test
```{r}
immport.b4 <- b4.r1[b4.r1$Data_Source == "ImmPort_DR19",]
imm.t <- sum(immport.b4$count)
unosopt.b4 <- b4.r1[b4.r1$Data_Source == "UNOS_OPTN",]
unos.t <- sum(unosopt.b4$count)
ucsfemr.b4 <- b4.r1[b4.r1$Data_Source == "UCSF_EMR",]
emr.t <- sum(ucsfemr.b4$count)
```

### within immport
```{r}
imm.c <- immport.b4[immport.b4$Race=="White_or_Caucasian",]
imm.nc <- immport.b4[immport.b4$Race!="White_or_Caucasian",]

a <- merge(imm.c, imm.nc, by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] = 0

prob.x <- a$count.x/imm.t
prob.y <- a$count.y/imm.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

```

#### merge immport and unos
```{r}
#### CAUCASIAN ####
a <- merge(unosopt.b4[unosopt.b4$Race=="White_or_Caucasian",], immport.b4[immport.b4$Race=="White_or_Caucasian",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] = 0

## KS test
prob.x <- a$count.x/unos.t
prob.y <- a$count.y/imm.t
ks.test(prob.x,prob.y, exact=TRUE, alternative="two.sided")

#### NON-CAUCASIAN ####
a <- merge(unosopt.b4[unosopt.b4$Race!="White_or_Caucasian",], immport.b4[immport.b4$Race!="White_or_Caucasian",], by="Donation_Age", all.x=TRUE, all.y=TRUE)
a <- a[,c("Donation_Age","count.x","count.y")]
a[is.na(a)] = 0

## KS test
prob.x <- a$count.x/unos.t
prob.y <- a$count.y/imm.t
ks.test(prob.x, prob.y, exact=TRUE, alternative="two.sided")

```


------ LOOKING ONLY AT IMMPORT RELATIONSHIP------
# input
```{r}
dm <- read.delim("/Users/jiemingchen/Documents/transplantation/a_donor/immport/finals.iTx/final.iTx.DM.v15.txt", header = T, sep = "\t") %>% 
          select(Sub_Org_Accession, AGE_TRANSPLANT, PHI_GENDER, DEMO_RACE_MERGED)
rp <- read.delim("/Users/jiemingchen/Documents/transplantation/a_donor/immport/finals.iTx/final.iTx.recipient.v15.txt", header = T, sep = "\t")
idata <- merge(dm, rp, by = "Sub_Org_Accession")
idata <- idata %>% 
            mutate(AGE_TRANSPLANT = round(AGE_TRANSPLANT)) %>%
            mutate(DEMO_DNR_RELATE_COMB = 
  ifelse(DEMO_DNR_RELATE == "Child", "Biological,_Child", 
  ifelse(DEMO_DNR_RELATE == "Biological,_Half_Sibling" | DEMO_DNR_RELATE == "Sister_(full_or_half)" | DEMO_DNR_RELATE == "Brother_(full_or_half)" | DEMO_DNR_RELATE == "Identical_Twin", "Biological,_Full_Sibling",
  ifelse(DEMO_DNR_RELATE == "Nephew" | DEMO_DNR_RELATE == "Niece" | DEMO_DNR_RELATE == "Aunt" | DEMO_DNR_RELATE == "Uncle" | DEMO_DNR_RELATE == "Grandfather" | DEMO_DNR_RELATE == "Grandmother" | DEMO_DNR_RELATE == "Cousin", "Biological,_Other_Relative_(Specify)",
  ifelse(DEMO_DNR_RELATE == "Father" | DEMO_DNR_RELATE == "Mother", "Biological,_Parent", 
  ifelse(DEMO_DNR_RELATE == "Unknown", NA, 
  ifelse(DEMO_DNR_RELATE == "Non-Biological,_Unrelated,_Directed_Dona" | DEMO_DNR_RELATE == "Non-Biological,_Unrelated,_Non-Directed" | DEMO_DNR_RELATE == "Non-Biological,_Unrelated_Paired_Exchang" | DEMO_DNR_RELATE == "Non-Biological,_Other_Unrelated_Directed", "Non-Biological,_Unrelated", as.character(DEMO_DNR_RELATE)))))))) %>%
  filter(DEMO_DNR_RELATE != "Living_Related" & DEMO_DNR_RELATE != "Living_related" & DEMO_DNR_RELATE != "Living_unrelated" & DEMO_DNR_RELATE != "Living_Unrelated")

## split at age 25
idata.lt25 <- idata[idata$AGE_TRANSPLANT <= 25,]
idata.mt25 <- idata[idata$AGE_TRANSPLANT > 25,]
```

# all subjects
```{r}
age.rship <- ddply(.data = idata, .variables = c("AGE_TRANSPLANT", "DEMO_DNR_RELATE_COMB"), .fun=summarise,
                   count = length(AGE_TRANSPLANT)) 
colnames(age.rship) <- c("AGE_TRANSPLANT", "DEMO_DNR_RELATE_COMB", "count")
age.rship <- age.rship %>% filter(!is.na(DEMO_DNR_RELATE_COMB))
```
## bar plot
```{r}
##-----------------------------------------------------------
## relationships
## BAR PLOT
pmain <- ggplot(age.rship, aes(AGE_TRANSPLANT, count, fill = DEMO_DNR_RELATE_COMB), xlim = c(0,80))
phisto <- geom_bar(stat = "identity")
plabels <- labs(x="Age",y="Count")
pticks <- scale_x_continuous(breaks=seq(min(age.rship$AGE_TRANSPLANT),max(age.rship$AGE_TRANSPLANT),by=1))
paxes <-  theme(axis.title.x = element_text(face = "bold",colour = "black", size = 20),
              axis.title.y = element_text(face = "bold",colour = "black", size = 20),
              axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15))
ptitle <- theme(plot.title = element_text(lineheight = 50, face = "bold", size = 30))
plegend <- theme(legend.title = element_text(face = "bold", size = 15), legend.text = element_text(size = 14))
pline <- geom_vline(xintercept=25, color="red")
## set1 usually has only 8 colors
getPalette <- colorRampPalette(brewer.pal(9, "Set1"))
## how many colors do we need
colorCount <- length(unique(age.rship$DEMO_DNR_RELATE_COMB))
## "manually" add colors instead of scale_fill_brewer fixed brwer colors
# newcolors <- c("#E41A1C","#66628D","#66628D","#419486","#5A9D5A","#66628D","#999999","#999999","black","#999999","#999999","#999999")
pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_fill_manual(values = getPalette(colorCount)) + pline
# pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_fill_manual(values = newcolors) + pline
# ggsave("age_reciprship_bar_newcolors.pdf", device = "pdf")
```

## line plot
```{r}
#
# LINE
pmain <- ggplot(age.rship, aes(AGE_TRANSPLANT, count, color = DEMO_DNR_RELATE_COMB), xlim = c(0,80))
phisto <- geom_line(size=1.5)
pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_color_manual(values = getPalette(colorCount)) + pline
# pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_color_manual(values = newcolors) + pline
# ggsave("age_reciprship_line_newcolors.pdf", device = "pdf")



```

## split relationships by gender using line plots
```{r}
age.rship.gender <- ddply(.data = idata, .variables = c("AGE_TRANSPLANT", "PHI_GENDER", "DEMO_DNR_RELATE_COMB"), .fun=summarise, count = length(AGE_TRANSPLANT))
age.rship.gender$label <- paste(age.rship.gender$PHI_GENDER,age.rship.gender$DEMO_DNR_RELATE_COMB,sep="-")
total.rship.gender <- ddply(.data = idata, .variables = c("PHI_GENDER", "DEMO_DNR_RELATE_COMB"), .fun=summarise, count = length(AGE_TRANSPLANT)) %>% filter(!is.na(DEMO_DNR_RELATE_COMB))
total.rship.gender$label <- paste(total.rship.gender$PHI_GENDER,total.rship.gender$DEMO_DNR_RELATE_COMB,sep="-")
age.rship.gender <- merge(age.rship.gender, total.rship.gender, by="label")
age.rship.gender <- age.rship.gender %>% filter(!is.na(DEMO_DNR_RELATE_COMB.x)) %>% mutate(proportion = count.x/count.y)

## LINE
pmain <- ggplot(age.rship.gender, aes(AGE_TRANSPLANT, count.x, color = DEMO_DNR_RELATE_COMB.x), xlim = c(0,80), ylim = c(0,70))
phisto <- geom_line(size=1.5)
pfacet <- facet_grid(PHI_GENDER.x ~., scale = "free_y")
pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_color_manual(values = getPalette(colorCount)) + pline + pfacet
# pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_color_manual(values = newcolors) + pline + pfacet
# ggsave("age_gender_split_reciprship_line_newcolors.pdf", device = "pdf")

pmain <- ggplot(age.rship.gender, aes(AGE_TRANSPLANT, proportion, color = DEMO_DNR_RELATE_COMB.x), xlim = c(0,80), ylim = c(0,70))
phisto <- geom_line(size=1.5)
pfacet <- facet_grid(~ PHI_GENDER.x, scale = "free_y")
pmain + phisto + plabels + pticks + paxes + ptitle + plegend +
  scale_color_manual(values = getPalette(colorCount)) + pline + pfacet
# #
# ## BAR
# 
# pmain <- ggplot(age.rship.gender, aes(AGE_TRANSPLANT, count, fill = DEMO_DNR_RELATE_COMB))
# phisto <- geom_bar(stat = "identity")
# pfacet <- facet_grid(PHI_GENDER ~., scale = "free_y")
# pmain + phisto + plabels + pticks + paxes + ptitle + plegend + pline + pfacet 
# # pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_fill_manual(values = newcolors) + pline + pfacet
# # ggsave("age_gender_split_reciprship_bar.pdf", device = "pdf")

```

## split relationship by ethnicity
```{r}
age.rship.eth <- ddply(.data = idata, .variables = c("AGE_TRANSPLANT", "DEMO_RACE_MERGED", "DEMO_DNR_RELATE_COMB"), .fun=summarise, count = length(AGE_TRANSPLANT), proportion = length(AGE_TRANSPLANT)/nrow(idata))
colnames(age.rship.eth) <- c("AGE_TRANSPLANT", "DEMO_RACE_MERGED", "DEMO_DNR_RELATE_COMB", "count", "proportion")
age.rship.eth <- age.rship.eth %>% filter(!is.na(DEMO_DNR_RELATE_COMB))

## LINE
pmain <- ggplot(age.rship.eth, aes(AGE_TRANSPLANT, proportion, color = DEMO_DNR_RELATE_COMB), xlim = c(0,80), ylim = c(0,70))
phisto <- geom_line(size=1.5)
pfacet <- facet_grid(~ DEMO_RACE_MERGED, scale = "free_y")
pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_color_manual(values = getPalette(colorCount)) + pline + pfacet
# pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_color_manual(values = newcolors) + pline + pfacet
# ggsave("age_gender_split_reciprship_line_newcolors.pdf", device = "pdf")

```

# relationship Trends in UNOS
```{r}
idata.unos <- 
  read.table("/Users/jiemingchen/Documents/transplantation/a_donor/unos/unos-LIVING_DONOR_DATA_DM_race_organ_relation-kidney-us-128407.rshipmod.txt", header = T, stringsAsFactors = F) 
age.rship.gender.unos <- ddply(.data = idata.unos, .variables = c("AGE_DON", "GENDER", "LIV_DON_TY_MERGED"), .fun=summarise, count = length(AGE_DON))
age.rship.gender.unos$label <- paste(age.rship.gender.unos$GENDER,age.rship.gender.unos$LIV_DON_TY_MERGED,sep="-")
total.rship.gender.unos <- ddply(.data = idata.unos, .variables = c("GENDER", "LIV_DON_TY_MERGED"), .fun=summarise, count = length(AGE_DON)) %>% filter(!is.na(LIV_DON_TY_MERGED))
total.rship.gender.unos$label <- paste(total.rship.gender.unos$GENDER,total.rship.gender.unos$LIV_DON_TY_MERGED,sep="-")
age.rship.gender.unos <- merge(age.rship.gender.unos, total.rship.gender.unos, by="label")
age.rship.gender.unos <- age.rship.gender.unos %>% filter(!is.na(LIV_DON_TY_MERGED.x)) %>% mutate(proportion = count.x/count.y)

## LINE
pmain <- ggplot(age.rship.gender.unos, aes(AGE_DON, count.x, color = LIV_DON_TY_MERGED.x), xlim = c(15,77), ylim = c(0,70))
phisto <- geom_line(size=1.5)
pfacet <- facet_grid(GENDER.x ~., scale = "free_y")
pmain + phisto + plabels + pticks + paxes + ptitle + plegend + scale_color_manual(values = getPalette(colorCount)) + pline + pfacet
```