---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

```{r}
library(animation)
library(visNetwork)
library(dplyr)
library(reshape2)
library(igraph)
library(data.table)
library(rgl)
library(ggplot2)

## input
outcomes <- read.table("/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/immport2cytoscape/txp_immport2cytoscape_seq-merged-relive01_02_03_immport2tsv-noControls_outcomes_kidney.txt", sep = "\t", header = T, stringsAsFactors = F)

## subsetting
# those with no time points, excluding those lines with NA
outcomes.notimepoints <- outcomes %>% 
                    filter(eventA_min == 5000000000 | eventB_min == 5000000000) %>%
                    filter(eventA_min != 5000000001 & eventB_min != 5000000001)
# those with time points
# this is my nw
outcomes.wtimepoints <- outcomes %>% 
                    filter(eventA_min != 5000000000 & eventB_min != 5000000000) %>%
                    filter(eventA_min != 5000000001 & eventB_min != 5000000001)
                    
```


# turning raw data to igraph object
# node and edge definitions
```{r}
## create networks from outcomes with timepoints
## also when time_min is >0 (posttx)
nw.prefiltered <- outcomes.wtimepoints[c(1,3,2,4,5,6,7)]
indivs.w.conditions.pretx <- nw.prefiltered %>% filter(eventA_min < 0 | eventB_min < 0)
nw <- nw.prefiltered %>% filter(eventA_min >= 0 & eventB_min >= 0)
names(nw) <- c("from","to","label","eventA_min","eventA_max","eventB_min","eventB_max")
write.table(nw, file = "/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/filter_input_table_for_network.txt", sep = "\t", quote = F, row.names = F)

## create network
graph <- graph.data.frame(nw, directed=T)
V(graph)$indegree <- centr_degree(graph, mode="in")$res

## nodes
nodes.labels <- read.table("/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/tx_LD_outcome_node_info.txt", header = T)
names(nodes.labels) <- c("id","label","definition")

nodes <- data.frame(id = unique(append(outcomes.wtimepoints$eventA, outcomes.wtimepoints$eventB)))
nodes <- merge(nodes, nodes.labels, by="id") 

nodes.indeg <- data.frame(id = V(graph)$name, in_degree = centr_degree(graph, mode="in")$res)
nodes <- merge(nodes, nodes.indeg, by="id") %>% 
            mutate(size = ifelse(in_degree <= 50, in_degree, 50))

nodes$group <- c(rep(1,4),2,rep(1,5),2,rep(1,4),2,rep(1,4),2,1,2,rep(1,14))
nodes$color <- c(rep(1,4),2,rep(1,5),2,rep(1,4),2,rep(1,4),2,1,2,rep(1,14))
nodes$shape <- "dot"
nodes$borderWidth <- 2 # Node border width

# nodes$color.background <- c("slategrey", "tomato", "gold")[nodes$media.type]
# nodes$color.border <- "black"
# nodes$color.highlight.background <- "orange"
# nodes$color.highlight.border <- "darkred"
# nodes <- get.data.frame(graph, what="vertices")
# nodes <- data.frame(id = nodes$name, title = nodes$name, group = nodes$indegree, indegree = nodes$indegree)
# setnames(nodes, "indegree", "in-degree centrality")
# nodes <- nodes[order(nodes$id, decreasing = F),]


## edges

# EDGES BY INDIV
edges.by.indiv <- get.data.frame(graph, what="edges")
edges.by.indiv <- edges.by.indiv %>% 
                    mutate(length_full = eventB_min - eventA_min, 
                           length = ifelse(length_full <= 1000, length_full, 1000))

# EDGES BY WIDTH - width of edges = number of indiv
edges.by.clust <- edges.by.indiv %>% 
                    group_by(from,to) %>% 
                    summarize(width=n())

#color
# colorCount = length(unique(edges.by.indiv$label))
# colors = colorRampPalette(brewer.pal(9, "Set1"))
# edges.by.indiv$color <- as.numeric(as.factor(edges.by.indiv$label))
# 

# edge by clustering -- where width = number of individuals
edges.by.clust <- edges.by.indiv %>%
                      group_by(from,to) %>% summarize(width_full=n()) %>%
                      mutate(width = ifelse(width_full <= 50, width_full, 50))

write.table(edges.by.clust, file = "/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/txp_immport2cytoscape.filtered.edgesbyclust.txt", sep = "\t", quote = F, row.names = F)


# EDGES BY MEAN TIME
edges.by.meantime <- edges.by.indiv %>% 
                        mutate(newname = paste(from,"-",to, sep="")) %>%
                        group_by(from,to) %>%
                        summarize(length_full_median = median(length_full),
                              length_full_mean = mean(length_full)) %>%
                        mutate(length = ifelse(length_full_mean <= 1000, length_full_mean, 1000))
                        
edges.by.meantime <- merge(edges.by.meantime, nodes.labels, by.x = "from", by.y = "id")
edges.by.meantime <- merge(edges.by.meantime, nodes.labels, by.x = "to", by.y = "id")

##### calculate centrality of network
# nw.betweenness <- as.data.frame(betweenness(graph))


```

# plot mean time vs node size from post-transplant (boxplot)
```{r}
o.ori <- read.table("/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/immport2cytoscape/txp_immport2outcomes-merged-relive01_02_03_immport2tsv-noControls_posttx_outcomes_kidney.txt", header = T, stringsAsFactors = F)


o <- o.ori %>%
        filter(!is.na(event_min) & !is.na(event_max)) %>%
        filter(event_min >= 0)

## add the tx node only on non-NA individuals
tx <- data.frame(subject_ID=unique(o$subject_ID), event="transplant",event_min=0,event_max=0)
o <- rbind(o, tx)

o <- merge(o, nodes.labels, by.x = "event", by.y = "id")


# plot violin plots
# x axis      = edgelength
# y axis      = label
# bubble size = number of subjects with that event
# edge length = mean time to event post transplant
# o %>%
#   ggplot(aes(x=event_min/365, y=label)) + 
#   geom_point() +
#   theme(legend.position = "none", 
#         axis.text.x = element_text(size = 14, color = "black"),
#         axis.text.y = element_text(size = 14)) + 
#   theme(legend.position="right") + 
#   # scale_fill_manual(values = colors, na.value = "white") +
#   # scale_color_manual(values = colors, na.value = "white") + 
#   labs(x="time from transplant (years)",y="post-donation outcomes") + 
#   scale_size_area(max_size = 14) +
#   theme(legend.title = element_text(size = 14, face="bold"), legend.text=element_text(size = 14)) +
#   theme(axis.title.y = element_text(size = rel(1.5))) +
#   theme(axis.title.x = element_text(size = rel(1.5)))
```

# plot mean time vs node size from post-transplant (bubble)
```{r}
# MEAN edge time
# node size = number of individuals
op <- o %>% 
        group_by(event) %>%
        summarize(nodesize = n(),
                  edgelength_mean = mean(event_min),
                  edgelength_median = median(event_min))
op <- merge(op, nodes.labels, by.y = "id", by.x = "event")
op <- op[order(op$edgelength_mean,-op$nodesize),]

op$label <- factor(op$label, levels=op[rev(order(op$edgelength_mean,-op$nodesize)),]$label)
op[op$event=="transplant", "nodesize"] <- 0 ## comment out this line to obtain size of transplant

op$label_new <- paste(op$label," (",op$nodesize,")", sep="")
op$label_new <- factor(op$label_new, levels=op[rev(order(op$edgelength_mean,op$nodesize)),]$label_new)

write.table(op, file = "/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/txp_immport2cytoscape.outcome.nodeinfo.txt", sep = "\t", quote = F, row.names = F)


# plot bubbles
# x axis      = edgelength
# y axis      = label
# bubble size = number of subjects with that event
# edge length = mean time to event post transplant
op %>%
  # ggplot(aes(x=edgelength_mean/365, y=label, size=nodesize)) + 
  ggplot(aes(x=edgelength_mean/365, y=label, size=nodesize)) + ## y labels with size
  # geom_point(shape = 23, fill = "black", color = "white") + ## diamonds not looking good
  geom_point(shape = 21, fill = "black", color = "white") + ## circles
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14)) + 
  geom_vline(xintercept = 2, color = "slategrey", linetype = "dashed") + 
  # scale_fill_manual(values = colors, na.value = "white") +
  # scale_color_manual(values = colors, na.value = "white") + 
  labs(x="time from transplant (years)",y="post-donation outcomes") + 
  scale_size(breaks=c(10,50,100,200,800,1000), 
             guide=guide_legend(title="living donors")) + ## legend with user-defined bubble sizes
  # scale_size_area(max_size = 14) +
  theme(legend.position="right") +
  theme(legend.position="none") + ## no legend
  theme(legend.title = element_text(size = 14, face="bold"), legend.text=element_text(size = 14)) +
  theme(axis.title.y = element_text(size = rel(1.5))) +
  theme(axis.title.x = element_text(size = rel(1.5))) + 
  scale_x_continuous(limits = c(0,45))

# # MEDIAN edge time ####
# op <- op[order(op$edgelength_median,-op$nodesize),]
# op$label <- factor(op$label, levels=op[rev(order(op$edgelength_median,-op$nodesize)),]$label)
# 
# op %>%
#   ggplot(aes(x=edgelength_median/365, y=label, size=nodesize)) + 
#   geom_point(shape = 21) +
#   theme(legend.position = "none", 
#         axis.text.x = element_text(size = 14, color = "black"),
#         axis.text.y = element_text(size = 14)) + 
#   theme(legend.position="right") + 
#   # scale_fill_manual(values = colors, na.value = "white") +
#   # scale_color_manual(values = colors, na.value = "white") + 
#   labs(x="time from transplant (years)",y="post-donation outcomes") + 
#   scale_size_area(max_size = 14) +
#   theme(legend.title = element_text(size = 14, face="bold"), legend.text=element_text(size = 14)) +
#   theme(axis.title.y = element_text(size = rel(1.5))) +
#   theme(axis.title.x = element_text(size = rel(1.5)))


```

# plot mean time vs node size from post-transplant 2 (boxplot) 
```{r}
# plot boxplots
# x axis      = edgelength
# y axis      = label
# bubble size = number of subjects with that event
# edge length = mean time to event post transplant

o <- merge(o, op) ## I want the edgelength whcih is the mean

## reorder the levels; by MEAN (not MEDIAN)
# o$label <- factor(o$label, levels = levels(factor(op$label, levels=op[rev(order(op$edgelength_median,-op$nodesize)),]$label)))
# o$label_new <- factor(o$label_new, levels = levels(factor(op$label, levels=op[rev(order(op$edgelength_median,-op$nodesize)),]$label_new)))
o$label <- factor(o$label, levels = levels(factor(op$label, levels=op[rev(order(op$edgelength_mean,op$nodesize)),]$label)))
o$label_new <- factor(o$label_new, levels = levels(factor(op$label, levels=op[rev(order(op$edgelength_mean,op$nodesize)),]$label_new)))

# only x can be factor, y has to be values
o %>%
  # ggplot(aes(y=event_min/365, x=label)) + 
  ggplot(aes(y=event_min/365, x=label_new)) +  # y labels with size
  geom_boxplot(aes(fill = "orange", color = "orange"), fatten = 0, width = 0.2) + ## fatten = 0 no median lines
  # geom_point() +
  # geom_violin() +
  geom_hline(yintercept = 2, color = "slategrey", linetype = "dashed") + 
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14)) + 
  theme(legend.position="right") + 
  # scale_fill_manual(values = colors, na.value = "white") +
  # scale_color_manual(values = colors, na.value = "white") + 
  labs(y="time from transplant (years)", x="post-donation outcomes") + 
  scale_size_area(max_size = 14) +
  theme(legend.position="none") + ## no legend
  # theme(legend.title = element_text(size = 14, face="bold"), legend.text=element_text(size = 14)) +
  theme(axis.title.y = element_text(size = rel(1.5))) +
  theme(axis.title.x = element_text(size = rel(1.5))) + 
  coord_flip()
  
``` 

# plotting with igraph
```{r}
## set attributes
## set indegree as vertex size
V(graph)$size <- ifelse(centr_degree(graph, mode="in")$res > 50, 50, centr_degree(graph, mode="in")$res) 

## plot
c <- layout_in_circle(graph)
s <- layout.spring(graph)

# doing mds on the network using edge length as attribute
e <- graph.edgelist(as.matrix(edges.by.indiv[,1:2]))
E(e)$length <- edges.by.indiv$length
adjm_edgelength <- get.adjacency(e, attr='length', sparse = F)
m <- layout_with_mds(graph, dist = adjm_edgelength)

# edges by meantime
e2 <- graph.edgelist(as.matrix(edges.by.meantime[,1:2]))
E(e2)$length <- edges.by.meantime$length
adjm_edgelength2 <- get.adjacency(e2, attr='length', sparse = F)
m2 <- layout_with_mds(graph, dist = adjm_edgelength2)

# xycoordinates from ds.js
xycoord <- read.table("/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/igraph/getCoordinates_in_graph_order.txt", header = F, stringsAsFactors = F)
# weights <- c(-500,seq(17,1),0,seq(1,17),-500) * 2
xycoord$V1 <- xycoord$V1 *10
xycoord$V2 <- xycoord$V2 *10

plot(graph, 
     edge.arrow.size=.4, 
     edge.curved=.1,
     edge.label=NA,
     arrow.mode = 2,
     layout = m2)
     # coords = xycoord,
     # rescale = F,
     # xlim = c(-1276, 1276),
     # ylim = c(-522,360)) 

## plot in interactive 2D graphics 
# tkplot(graph, 
#      edge.arrow.size=.4, 
#      edge.curved=.1,
#      edge.label=NA,
#      arrow.mode = 2,
#      layout = m,
#      coords = xycoord,
#      rescale = F,
#      xlim = c(-1276, 1276),
#      ylim = c(-522,360))

# all possible layouts in igraph
# layouts <- grep("^layout_", ls("package:igraph"), value=TRUE)[-1] 
# layouts <- layouts[!grepl("bipartite|sugiyama", layouts)] # Remove layouts that do not apply to our graph.
# par(mfrow=c(3,3), mar=c(1,1,1,1))
# for (layout in layouts) {
#   print(layout)
#   l <- do.call(layout, list(graph)) 
#   plot(graph, 
#        edge.arrow.size=.4, 
#        edge.label=NA, 
#        edge.curved=.1, 
#        arrow.mode=2, 
#        layout=l, 
#        main=layout, 
#        dist = adjm_edgelength) 
# }

```

# plotting interactive visualization with visNetwork: trajectory
## 1) length is equal to accurate length of time from event A to B
```{r}

## to hide labels you have to change the column name
edges.by.indiv <- edges.by.indiv %>% mutate(label_hide = label) 
edges.by.indiv$label <- '' 


visNetwork(nodes, edges.by.indiv) %>%
  visOptions(selectedBy = "group", highlightNearest = TRUE, nodesIdSelection = TRUE) %>% # menu for nodes
  visPhysics(stabilization = FALSE) %>% 
  visNodes(font = list(size = 30)) %>%
  visEdges(arrows = "to") %>% ## hide labels by replacing column name "label" to something else
  visHierarchicalLayout(direction = "LR")  %>% # same as visLayout(hierarchical = TRUE)  
  visInteraction(dragNodes = TRUE, navigationButtons = TRUE,
                 dragView = TRUE, 
                 zoomView = TRUE) 

# visSave(g, file="/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/igraph/test.html", selfcontained = TRUE, background = "white")
```

## 2) edges are by mean time
```{r}

## to hide labels you have to change the column name
# edges.by.meantime <- edges.by.meantime
# edges.by.meantime$label <- '' 


visNetwork(nodes, edges.by.meantime) %>%
  visOptions(selectedBy = "group", highlightNearest = TRUE, nodesIdSelection = TRUE) %>% # menu for nodes
  visPhysics(stabilization = FALSE) %>% 
  visNodes(font = list(size = 30)) %>%
  visEdges(arrows = "to") %>% ## hide labels by replacing column name "label" to something else
  visHierarchicalLayout(direction = "LR")  %>% # same as visLayout(hierarchical = TRUE)  
  visInteraction(dragNodes = TRUE, navigationButtons = TRUE,
                 dragView = TRUE, 
                 zoomView = TRUE) 

# visSave(g, file="/Users/jiemingchen/Documents/transplantation/a_donor/immport/network_traj_analyses/igraph/test.html", selfcontained = TRUE, background = "white")
```

## 3) length is hidden = length does not mean anything
```{r}
edges.by.indiv2 <- edges.by.indiv %>% 
  rename(length_hide = length)

visNetwork(nodes, edges.by.indiv2) %>%
  visOptions(selectedBy = "group", highlightNearest = TRUE, nodesIdSelection = TRUE) %>% # menu for nodes
  visPhysics(stabilization = FALSE) %>% 
  visNodes(font = list(size = 30)) %>%
  visEdges(arrows = "to") %>% 
  visHierarchicalLayout(direction = "LR")  %>% # same as visLayout(hierarchical = TRUE)  
  visInteraction(dragNodes = TRUE, navigationButtons = TRUE,
                 dragView = TRUE, 
                 zoomView = TRUE)

```

## 4) edges by cluster -- where width = number of individuals
```{r}

visNetwork(nodes, edges.by.clust) %>%
  visOptions(selectedBy = "group", highlightNearest = TRUE, nodesIdSelection = TRUE) %>% # menu for nodes
  visPhysics(stabilization = FALSE) %>% 
  visNodes(font = list(size = 30)) %>%
  visEdges(arrows = "to") %>% 
  visHierarchicalLayout(direction = "LR")  %>% # same as visLayout(hierarchical = TRUE)  
  # visIgraphLayout(layout = "layout_in_circle")  %>%
  visInteraction(dragNodes = TRUE, navigationButtons = TRUE,
                 dragView = TRUE, 
                 zoomView = TRUE)

visNetwork(nodes, edges.by.clust) %>%
  visOptions(selectedBy = "in_degree", highlightNearest = TRUE, nodesIdSelection = TRUE) %>%
  visPhysics(stabilization = FALSE) %>% 
  visEdges(arrows = "to") %>% 
  visHierarchicalLayout(direction = "LR") # same as   visLayout(hierarchical = TRUE)

```

# traversing one subject: wip
```{r}
## traverse one subject: SUB130359
visNetwork(nodes, edges.by.indiv) %>%
  visOptions(selectedBy = "group", highlightNearest = TRUE, nodesIdSelection = TRUE) %>% # menu for nodes
  visPhysics(stabilization = FALSE) %>% 
  visNodes(font = list(size = 30)) %>%
  visEdges(arrows = "to", label = F) %>% 
  visHierarchicalLayout(direction = "LR")  %>% # same as visLayout(hierarchical = TRUE)  
  visInteraction(dragNodes = TRUE, navigationButtons = TRUE,
                 dragView = TRUE, 
                 zoomView = TRUE)
```

# histograms of edge and node distribution to set threshold to cut max 
```{r}

## edge length distribution
hist(edges.by.indiv$length)
hist(edges.by.indiv[edges.by.indiv$length < 1000,]$length)
edges.by.indiv <- edges.by.indiv %>% mutate(length_mod = ifelse(length <= 1000, length, 1000))

## node size distribution
hist(nodes$in_degree)
hist(nodes[nodes$in_degree < 100,]$in_degree)

## edge width distribution
hist(edges.by.clust$width)
```

# just a hairball
```{r}

# visNetwork(nodes,edges, width="100%", height="400px", main="tx_LD_outcomes")
# nodes
# nodes <- as.data.frame(unique(melt(cbind(outcomes.wtimepoints$eventA, outcomes.wtimepoints$eventB))$value))
# names(nodes) <- c('id')
# 
# # edges
# edges <- outcomes.wtimepoints[c(1,3,2,4,5,6,7)]
# names(edges) <- c("from","to","label","eventA_min","eventA_max","eventB_min","eventB_max")
# nodes$shape <- "dot"  
# nodes$shadow <- TRUE # Nodes will drop shadow
# nodes$title <- nodes$media # Text on click
# nodes$label <- nodes$type.label # Node label
# nodes$size <- nodes$audience.size # Node size
# nodes$borderWidth <- 2 # Node border width
# 
# # nodes$color.background <- c("slategrey", "tomato", "gold")[nodes$media.type]
# nodes$color.border <- "black"
# nodes$color.highlight.background <- "orange"
# nodes$color.highlight.border <- "darkred"
# 
# visNetwork(nodes, edges) 
```